---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

An introduction of the data and a description of the trends/books/items you are choosing to analyze (and why!)

Sarah J. Maas is a internationally acclaimed author, known for writing the Throne of Glass and A Court of Thorns and Roses series. According to Goodreads, her books have sold more than twelve million copies worldwide. I started reading her books in middle school and they are some of my most favorite reads; I’ve read both previously mentioned series over and over, hoping to relive the immersive experience. I wanted to explore how Sarah J. Maas’s books have fluctuated in terms of popularity over the years. I was extremely curious whether they have increased or decreased in popularity since I read the books. To conduct this analysis, I will use data from the Seattle Public Library’s open data portal, specifically the data on all of books written by Sarah J. Maas.

### Summary Information

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("summary.R")}
```

To gain a general understanding of the data, I found some summary values. All of the data is in relation to books written by Sarah J. Maas and checked out at the Seattle Public Library. The year with the lowest number of Sarah J. Maas checkouts is `r year_least_checkouts` while the year with the highest number of checkouts is `r year_most_checkouts`. I was also interested in which type of material is most and least frequently checked out. `r type_highest_avg_checkouts` is the material type with the highest average checkouts while `r type_lowest_avg_checkouts` is the material type with the lowest average checkouts. The series with the most total checkouts is the `r series_most_checkouts` while the series with the least total checkouts is the `r series_least_checkouts`. Similarly, the book with the most checkouts is `r book_most_checkouts` while the book with the least checkouts is `r book_least_checkouts`.

### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart_1.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart_2.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart_3.R")}
```


